Cluster Difference Imaging Photometric Survey. III. An Open Cluster Spanning 500\,Parsecs

### 2021-04-06 cluster_paper/ms.tex
update ms

\subsubsection{Cluster Star Sample}
\label{subsubsec:cluster}

@@

@@

\subsubsection{Kinematics $\otimes$ Rotation}

@@

Given the rotation periods from TESS, we can explore how far away from
the core we observe cluster members from the core in position and
velocity space.  To do this, we cross-matched the stars with detected
rotation periods against our original target list of \nkinematic\ Gaia
DR2 kinematic candidate members.  For a fair comparison, we considered
only the stars with $0.5<(Bp-Rp)_0<1.2$ for which our TESS pipeline
successfully made detrended light curves.  In other words---the base
sample is the stars for which we could have plausibly detected a
rotation period.  Faint stars with $Rp>16$ were beyond our light curve
selection limit, and stars not meeting the crowding requirements
previously described are omitted.  We selected the color range above
based on Figure~\ref{fig:rot}, since the density of period detections
seems to decrease once $(Bp-Rp)_0 \gtrsim 1.25$, {\it i.e.}, for
spectral types later than $\approx$K4V.

% IN POSITION
% Bin width: 25
% Rotators: [361  82  43  30  16  18  14   8   6   2   4   0   0   0   0
% 0   0   0 0   0]
% Comparison: [495 116  65  46  28  20  22  12  11   8   5   1   0   1
% 0   2   0   0 0   0]
% Rotators < 25pc : 0.72929 +/- 0.00013
% Rotators > 25pc : 0.66172 +/- 0.00023
%

The results are shown in Figure~\ref{fig:physical_x_rotn}.
Appendix~\ref{app:gaia6d_x_rotn} includes a supplemental visualization
in the observed space of Gaia measurements, rather than Cartesian
coordinates and tangential velocities.  The observed proper motions
however suffer from projection effects, discussed in
Appendix~\ref{app:vproj} {\bf TODO: merge these appendices}.  These
effects are important in the equatorial coordinate system due to the
proximity of some \cn\ members to the southern ecliptic pole.  For a
star at position $(\alpha, \delta)$, we correct for the projection
effect by comparing the observed proper motion $(\mu_{\alpha'},
\mu_\delta)$ to what the proper motion at the star's position would
have been if the star were comoving with the core of \cn.  This yields
a quantity we denote $\Delta \mu^{*}$, following
\citetalias{meingast_2021}.  We then convert these proper motion
differences to physical units by dividing by the distance to each
star, taken as the inverse parallax.

{\bf FIXME FIXME FIXME} 
The results suggest that the
fraction of kinematic members with TESS-detected rotation periods may
decrease with increasing distance from the halo core---both in
position and in velocity space.

%The inverse parallax approximation does not introduce significant
%biases for this sample, since the mean $\pi/\sigma_\pi$ and its
%standard deviation are 122 and 40 respectively; the minimum is 19.5.

We quantified this by computing the separation of the positions and 2D
velocities from the median \citetalias{cantatgaudin_gaia_2018} core
member {\bf TODO: clarify binning choice}.  The result is shown on the right side of
Figure~\ref{fig:physical_x_rotn}.  The halo extends to separations of
about 200\,pc in physical space from the cluster core.  The most
widely separated F2V-K2V stars with rotation periods (blue points in
Figure~\ref{fig:physical_x_rotn}) are separated by $\approx$430\,pc.
The total length of the structure is therefore 400--500\,pc, depending
on which members of the halo are chosen as the ``tips'' on either end.
Broadly speaking, this agrees with the structure of the halo reported
by \citetalias{kounkel_untangling_2019}.

{\bf FIXME update this text, make it less crappy... ditto the figure}
In projection-corrected tangential velocity space, the fraction of
stars with rotation period detections remains high out to roughly
5\kms.  \citet{meingast_2021} by comparison required a physically
motivated cut in tangential velocity space of 1.5\kms.  Our results
show that at the expense of higher field star contamination rates,
bonafide members can be identified even out to higher velocity
separations.  {\bf todo: anything else to mention? e.g., the actual
numbers?}

@@

\section{Kinematics $\otimes$ Rotation}
\label{app:gaia6d_x_rotn}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.95\textwidth]{f4.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Gaia-based components of NGC\,2516 in position and
  velocity space, cross-matched against the rotators.} The plotted
  stars are those with $0.5<(Bp-Rp)_0<1.2$ that meet the crowding
  restrictions described in Section~\ref{subsec:tess}: they should
  have been sufficiently bright and non-crowded to enable rotation
  period detection.  We caution that the appearance of fewer
  non-rotators being present toward the core is due the layering of
  the plot: quantitatively, stars toward the cluster center do not all
  show rotation periods in our analysis (see
  Figure~\ref{fig:hist_physical_x_rotn}).
  \label{fig:gaia6d_x_rotn}
	}
\end{figure*}

Figure~\ref{fig:gaia6d_x_rotn} is an alternative visualization of the
data shown in Figure~\ref{fig:physical_x_rotn}.  The rotation periods
in this diagram correspond to ``Set $\mathcal{A}$'' as described in
Section~\ref{subsec:tess}.  Many stars in the outer region of the
cluster show rotation periods consistent with a young age.  One issue
with the visualization as displayed however is that it hides the
amount of non-rotators toward the cluster center: we did not detect
rotation periods for roughly one in four stars at the cluster
center (see Figures~\ref{fig:physical_x_rotn}
and~\ref{fig:hist_physical_x_rotn}).
% 
% 
% 
% \section{Rotation $\otimes$ RUWE}
% \label{app:ruwe}
% 
% \begin{figure*}[t]
% 	\begin{center}
% 		\leavevmode
% 		\includegraphics[width=0.95\textwidth]{f6.pdf}
% 	\end{center}
% 	\vspace{-0.7cm}
% 	\caption{ {\bf Rotation versus color, colored by RUWE.}
% 		Looks like on the slow sequence, there's more yellow at the bottom?
% 		\label{fig:rotn_X_RUWE}
% 	}
% \end{figure*}
% 
% Figure~\ref{fig:rotn_X_RUWE}.
% 
% 
% \section{Kinematics $\otimes$ Rotation in EDR3}
% 
% \begin{figure*}[t]
% 	\begin{center}
% 		\leavevmode
% 		\includegraphics[width=0.95\textwidth]{f7.pdf}
% 	\end{center}
% 	\vspace{-0.7cm}
% 	\caption{ {\bf Gaia-based components of NGC\,2516 in position and
%     velocity space, cross-matched against the rotators. Analog of
%     Figure~\ref{fig:gaia6d_x_rotn}, but showing EDR3 kinematics.}
% 		\label{fig:gaia6d_x_rotn_EDR3}
% 	}
% \end{figure*}
% 
% How did we do?
% Figure~\ref{fig:gaia6d_x_rotn_EDR3} lets us check with a more precise
% astrometric solution.

>>>>

\subsubsection{Methodology}
\label{subsubsec:cluster}

@@

\subsubsection{Rotation-Color Diagrams}

@@

\subsubsection{Comparing Kinematics and Rotation}

@@

The TESS rotation periods can provide an independent check on the
Gaia-derived kinematic cluster memberships.  We explored this by
cross-matching the stars with detected TESS rotation periods against
our original target list of \nkinematic\ Gaia DR2 members.

To visualize the results, we again opted for Cartesian coordinates,
and supplemented them by calculating the tangential stellar velocities
relative to the cluster center.  Appendix~\ref{app:vproj} discusses
the projection effect correction required for calculating the
tangential velocities: for a star at position $(\alpha, \delta)$, we
compare the observed proper motion $(\mu_{\alpha'}, \mu_\delta)$ with
what the proper motion at the star's position would have been if the
star were comoving with the core of \cn.  This yields a quantity we
denote $\Delta v^{*}$, per \citetalias{meingast_2021}.  We convert
these proper motion differences to physical units by dividing by the
distance to each star, taken as the inverse parallax.

The results are shown in the top two rows of
Figure~\ref{fig:physical_x_rotn}.  In these figures, to ensure a fair
comparison, we only show stars with $0.5<(Bp-Rp)_0<1.2$ for which our
TESS pipeline succeeding in making light curves.  In other words---the
base sample is the stars for which we could have plausibly detected a
rotation period.  This color range is preferred since the the slow
sequence becomes less defined for spectral types later than
$\approx$K4V.  The bins in the histograms (top-right) were chosen to
ensure that the same number of stars were in each bin.  Rotation
periods are detected for 66\% of stars at least 25\,pc from the
cluster core.  The average detection rate inside of 25\,pc, 73\%, is
slightly higher.

The lower panels of Figure~\ref{fig:physical_x_rotn} give a second
view of how field star contamination affects the outskirts of the
cluster.  Of the outermost 100 stars in position space with detected
rotation periods, $\approx$8 appear as outliers above the slow
sequence, compared to just a few for the innermost 100 stars.  For the
outermost 100 stars in velocity space, $\approx$13 appear as outliers.
After accounting for the stars that were expected to show rotation
periods and did not, this suggests field contaminant rates of
$\approx$40-50\% for the outermost cluster members in our target list.

Speaking in the positive though, our rotation period measurements show
that the halo extends to separations of $\approx$250\,pc in physical
space from the cluster core.  The most widely separated F2V-K2V stars
with rotation periods (blue points in
Figure~\ref{fig:physical_x_rotn}) are separated by $\approx$430\,pc.
The total length of the structure is therefore 400--500\,pc, depending
on which members of the halo are chosen as the ``tips'' on either end.
This agrees with the overall structure of the halo reported by
\citetalias{kounkel_untangling_2019}, with some minor caveats visible
in the top-left panels of Figure~\ref{fig:physical_x_rotn}.

In projection-corrected tangential velocity space, the fraction of
stars with rotation period detections remains high out to roughly
5\kms.  \citet{meingast_2021} by comparison required a physically
motivated cut in tangential velocity space of 1.5\kms.  Our results
show that at the expense of higher field star contamination rates,
bonafide members can be identified even out at higher velocity
separations.

@@

% \section{Kinematics $\otimes$ Rotation}
% \label{app:gaia6d_x_rotn}
% 
% \begin{figure*}[t]
% 	\begin{center}
% 		\leavevmode
% 		\includegraphics[width=0.95\textwidth]{f4.pdf}
% 	\end{center}
% 	\vspace{-0.7cm}
%   \caption{ {\bf Gaia-based components of NGC\,2516 in position and
%   velocity space, cross-matched against the rotators.} The plotted
%   stars are those with $0.5<(Bp-Rp)_0<1.2$ that meet the crowding
%   restrictions described in Section~\ref{subsec:tess}: they should
%   have been sufficiently bright and non-crowded to enable rotation
%   period detection.  We caution that the appearance of fewer
%   non-rotators being present toward the core is due the layering of
%   the plot: quantitatively, stars toward the cluster center do not all
%   show rotation periods in our analysis (see
%   Figure~\ref{fig:hist_physical_x_rotn}).
%   \label{fig:gaia6d_x_rotn}
% 	}
% \end{figure*}
% 
% Figure~\ref{fig:gaia6d_x_rotn} is an alternative visualization of the
% data shown in Figure~\ref{fig:physical_x_rotn}.  The rotation periods
% in this diagram correspond to ``Set $\mathcal{A}$'' as described in
% Section~\ref{subsec:tess}.  Many stars in the outer region of the
% cluster show rotation periods consistent with a young age.  One issue
% with the visualization as displayed however is that it hides the
% amount of non-rotators toward the cluster center: we did not detect
% rotation periods for roughly one in four stars at the cluster
% center (see Figures~\ref{fig:physical_x_rotn}
% and~\ref{fig:hist_physical_x_rotn}).
% 
% 
% 
% \section{Rotation $\otimes$ RUWE}
% \label{app:ruwe}
% 
% \begin{figure*}[t]
% 	\begin{center}
% 		\leavevmode
% 		\includegraphics[width=0.95\textwidth]{f6.pdf}
% 	\end{center}
% 	\vspace{-0.7cm}
% 	\caption{ {\bf Rotation versus color, colored by RUWE.}
% 		Looks like on the slow sequence, there's more yellow at the bottom?
% 		\label{fig:rotn_X_RUWE}
% 	}
% \end{figure*}
% 
% Figure~\ref{fig:rotn_X_RUWE}.
% 
% 
% \section{Kinematics $\otimes$ Rotation in EDR3}
% 
% \begin{figure*}[t]
% 	\begin{center}
% 		\leavevmode
% 		\includegraphics[width=0.95\textwidth]{f7.pdf}
% 	\end{center}
% 	\vspace{-0.7cm}
% 	\caption{ {\bf Gaia-based components of NGC\,2516 in position and
%     velocity space, cross-matched against the rotators. Analog of
%     Figure~\ref{fig:gaia6d_x_rotn}, but showing EDR3 kinematics.}
% 		\label{fig:gaia6d_x_rotn_EDR3}
% 	}
% \end{figure*}
% 
% How did we do?
% Figure~\ref{fig:gaia6d_x_rotn_EDR3} lets us check with a more precise
% astrometric solution.
